Extracting Angular Observables without a Likelihood\\and Applications to Rare Decays

### 2015-01-22 draft.tex
[draft] corrections on intro

The initial motivation for studying what we wish to call the \emph{Method of Moments}
is the determination of angular observables in the rare decay FCNC-mediated decay
$\bar{B}\to \bar{K}^*(\to \bar{K}\pi)\ell^+\ell^-$. However, we emphasize that
the method we describe in this letter is general, and it applies to arbitrary
decay or scattering processes. We find a previous work \cite{Dighe:1998vk}
advocates this method chiefly for the determination of angular observables in non-leptonic
$B$ decays, but also mentions the applicability to semileptonic decays. We improve
upon this previous work by studying correlations between the angular observables,
and sources of systematic uncertainties. Moreover, we will show that the method
at hand has two benefits over the usual approach based on likelihood fits. To wit:
\begin{enumerate}
    \item Likelihood fits have convergence problems for a small number of
        events, and can require reparametrizations and/or approximations
        for a successful fit to the signal PDF, see e.g. \cite{Egede:2008uy}.
        The method of moments has no such convergence problems.
    \item The results of likelihood fits suffer from systematic uncertainties
        and systematic bias (see e.g. \cite{Egede:2008uy}), especially in but not limited
        to cases in which the PDFs do not faithfully model the underlying physics. We will
        show that the method of moments does not suffer a very particular type of
        mismodelling that arises from introducing a cutoff in a partial wave expansion.
    \item Using the method of moments, the joint probability distribution of the
        angular observables converges towards a multivariate Gaussian distribution.
        This allows an easy transfer of correlation information from the experiments to
        interested theorists.
\end{enumerate}

@@

The determination of a detector's performance to detect signal events
with accurate determination of the event's angles is generally a difficult
task. In the ideal case, one would have an explicit probabilistic model of
the detector acceptance effects and could thus write down the full
forward model from which the measured events arise. In practice, this
would require much more CPU time than available, and one is therefore
forced to simplify the model. The standard approach is to generate
the \emph{true} particle events $\lbrace x_\text{t}^{(n)}\rbrace =
\lbrace(\vec{\nu}^{(n)}_\text{t},\vec\theta^{(n)}_\text{t})\rbrace$, $n=1,\dots,N_\text{t}$, from a PDF
assumed to describe the bare physical process, and to propagate those
particles through a detailed simulation of the detector.
The observable traces that the particles leave in the detector are fed
into reconstruction algorithms resulting in the \emph{detector} events
$\lbrace x^{(n)}_\text{d}\rbrace = \lbrace(\vec{\nu}^{(n)}_\text{d},
\vec\theta^{(n)}_\text{d})\rbrace$, $n=1,\dots,N_\text{d}$.
Most generally, the detected events $\lbrace x_\text{d}^{(k)}\rbrace$
are distributed as $P_\text{d}$,
\begin{equation}
    P_\text{d}(x_\text{d}) = \int P_\text{t}(x_\text{t}) E(x_\text{d}, x_\text{t}) \rmdx{x_\text{t}}\,.
\end{equation}
Here $P_\text{t}$ is the probability distribution of the true events, and the kernel $E(x_\text{d}, x_\text{t})$
parametrizes acceptance and resolution
effects. Note that $P_\text{d}$ is \emph{not} a PDF, since it is not normalized,
\begin{equation}
    \int P_\text{d}(x) \rmdx{x} \neq 1\,.
\end{equation}
For the sake of clarity, we will restrict ourselves
to pure acceptance effects and suppress resolution effects in the remainder of this section; i.e.,
\begin{equation}
    E(x_\text{d}, x_\text{t}) \equiv \eps(x_\text{t}) \delta(x_\text{d}, x_\text{t})\,.
\end{equation}
We will refer to $\eps$ as the \emph{detector acceptance} function. In what follows,
we propose a systematic methods to unfold all effects of $\eps$ through MC simulations
and the method of moments.\\

>>>>

The initial motivation for studying what we wish to call the
\emph{method of moments} is the determination of angular observables
in the rare decay FCNC-mediated decay $\bar{B}\to \bar{K}^*(\to
\bar{K}\pi)\ell^+\ell^-$. However, we emphasize that the method we
describe in this letter is general, and it applies to arbitrary decay
or scattering processes. We find a previous work \cite{Dighe:1998vk}
advocates this method chiefly for the determination of angular
observables in non-leptonic $B$ decays, but also mentions the
applicability to semileptonic decays. We improve upon this previous
work by studying correlations \fred{really? because of higher $l$}
between the angular observables, and by showing explicitly how to
incorporate uncertainties due to acceptance effects. Moreover, we show
that the method of moments has two major advantages over the usual
approach based on likelihood fits:
\begin{enumerate}
    \item Likelihood fits have convergence problems for a small number of
        events, and can require reparametrizations and/or approximations
        for a successful fit to the signal PDF, see e.g. \cite{Egede:2008uy}.
        The method of moments has no such convergence problems.
    \item The results of likelihood fits suffer from systematic uncertainties
        and systematic bias (see e.g. \cite{Egede:2008uy}), especially in but not limited
        to cases in which the PDFs do not faithfully model the underlying physics. We will
        show that the method of moments does not suffer a very particular type of
        mismodelling that arises from introducing a cutoff in a partial wave expansion.
    \item Using the method of moments, the joint probability distribution of the
        angular observables converges towards a multivariate Gaussian distribution.
        This allows an easy transfer of correlation information from the experiments to
        interested theorists.
\end{enumerate}

@@

The determination of a detector's performance to detect signal events
with accurate determination of the event's angles is generally a
difficult task. In the ideal case, one would have an explicit
probabilistic model of the detector acceptance effects and could thus
write down the full forward model from which the measured events
arise. In practice, this would require much more CPU time than
available, and one is therefore forced to simplify the model. The
standard approach is to generate the \emph{true} particle events
$\lbrace x_\text{t}^{(n)}\rbrace =
\lbrace(\vec{\nu}^{(n)}_\text{t},\vec\theta^{(n)}_\text{t})\rbrace$,
$n=1,\dots,N_\text{t}$, from a PDF assumed to describe the bare
physical process, and to propagate those particles through a detailed
simulation of the detector.  The observable traces that the particles
leave in the detector are fed into reconstruction algorithms resulting
in the \emph{detector} events $\lbrace x^{(n)}_\text{d}\rbrace =
\lbrace(\vec{\nu}^{(n)}_\text{d}, \vec\theta^{(n)}_\text{d})\rbrace$,
$n=1,\dots,N_\text{d}$.  In general, the distribution of the detected
events is
\begin{equation}
    P_\text{d}(x_\text{d}) = \int P_\text{t}(x_\text{t}) E(x_\text{d} | x_\text{t}) \rmdx{x_\text{t}}\,.
\end{equation}
Here $P_\text{t}$ is the probability distribution of the true events,
and the kernel $E(x_\text{d}| x_\text{t})$ parametrizes acceptance and
resolution effects. Note that $P_\text{d}$ is \emph{not} a PDF, since
it is not normalized,
\begin{equation}
    \int P_\text{d}(x) \rmdx{x} \neq 1\,.
\end{equation}
For the sake of clarity, we will restrict ourselves
to pure acceptance effects and suppress resolution effects in the remainder of this section; i.e.,
\begin{equation}
    E(x_\text{d}, x_\text{t}) \equiv \eps(x_\text{t}) \delta(x_\text{d}, x_\text{t})\,.
\end{equation}
We will refer to $\eps$ as the \emph{detector acceptance} function. In what follows,
we propose a systematic methods to unfold all effects of $\eps$ through MC simulations
and the method of moments.\\
